---
title: "PROCESSING STEPS OF THE DATABASE INCLUDING DESCRIPTIVE STATISTICS, CORRELATION ANALYSIS AND MODEL-BASED RANKING OF AD PARAMETERS EXTRACTED FROM THE LITERATURE"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Download packages

```{r}

library(psych)
library(visdat)
library(stats)
library(Hmisc)
library(ggplot2) 
library(ggpubr)
library(corrplot)
library(factoextra)
library(FactoMineR)
library(party)
library(gridExtra)
library(mlbench)
library(forecast)
library(GGally)
library(psych)
library(randomForestExplainer)
library(randomForest)
library(DALEX)
library(pdp)
library(dplyr)

```

# 2. Import the database
## 2.1. Raw database

```{r}
file1 <- 'rawalldata.csv'
database1 <- read.csv(file1, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
View(database1)
str(database1)

```
## 2.2. Qualitative variables converted

```{r}
file2 <- 'rawalldata_allquan.csv'
database2 <- read.csv(file2, sep = ";",dec = ",", na.strings = c("", " ", "NA"),stringsAsFactors = T)
View(database2)
str(database2)
```
## 2.3. Compelete database
```{r}
file3 <- 'Completedata.csv'
database3 <- read.csv(file3, sep = ";",dec = ",",na.strings = c("", " ", "NA"),stringsAsFactors = T)
View(database3)
str(database3)
```

# 3. Descriptive statistics of the whole database

```{r}
quantitative <- database1 [,c(1,2,3,5,6,10:12,14,15,17:26,28,29)]
qualitative <- database1 [,c(1,4,7:9,13,16,27)]
describe(quantitative [,-c(1)])
summary(qualitative)

```

## 3.1. Probability density of the start-up duration

```{r}
Meanduration <- mean(database3$Period_t)
plot(density(database3$Period_t), bty = "n", xlim= c(10,1000), yaxt="n", ylab ="", ylim=c(0,.01), lwd =2, xlab = "Total duration of start-up", main="Ditribution of start-up duration")
polygon(density(database3$Period_t), col="#FF66B2")
abline(v=mean(database3$Period_t), lwd =2, col="#330019")
abline(v=median(database3$Period_t), lwd=2, lty=3, col="#330019" )
text(x = mean(database3$Period_t) * 1.7,                   # Add text for mean
     y = mean(database3$Period_t) * 1.7,
     paste("Mean =", mean(database3$Period_t)),
     col = "red",
     cex = 2)
```

## 3.2. Plot the missingness of the database

```{r}
visdat::vis_miss(database1[,-1])
```

# 4. ACP analysis
## 4.1. Extraction of quantitative variables

```{r}
scaled_data <- scale(database3[,-c(1:4,7,10,16)], center = TRUE, scale = TRUE)
```

## 4.2. ACP of quantitative data

```{r}
PCA_database <- PCA(scaled_data, graph = FALSE)
```

## 4.3. eigenvalues

```{r}
fviz_eig(PCA_database, addlabels = TRUE, ylim = c(0, 50))

```

## 4.4. loading plot 

```{r}
plot1 <- fviz_pca_var(PCA_database, col.var = "cos2", axes = 1:2, 
             gradient.cols = c("blue", "yellow", "red"), 
             repel = TRUE, cex=2)
```

## 4.5. score plot 

```{r}
plot2 <- fviz_pca_ind(PCA_database, pointsize = database3$Period_t, 
             axes = 1:2, geom = c("text", "point"), 
             pointshape = 21, fill = "#E7B800",
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             legend.title = "Groups" , repel = T)

```

## 4.6. Plot the loading and the score plot together

```{r}
ggarrange(plot1, plot2, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)

```

# 5. Correlation analysis
## 5.1. Variable normality tests

```{r}
shapiro.test(as.numeric(database2$Ino_DG)) 
ggqqplot(as.numeric(database2$Ino_DG), ylab = "inoculum_digestate")
shapiro.test(as.numeric(database2$Ino_SS)) 
ggqqplot(as.numeric(database2$Ino_SS), ylab = "inoculum_SS")
shapiro.test(as.numeric(database2$Ino_Mix)) 
ggqqplot(as.numeric(database2$Ino_Mix), ylab = "inoculum_mixture")
shapiro.test(as.numeric(database2$Ino_CM)) 
ggqqplot(as.numeric(database2$Ino_CM), ylab = "inoculum_manure")
shapiro.test(as.numeric(database2$Incu)) 
ggqqplot(as.numeric(database2$Incu), ylab = "incubation")
shapiro.test(as.numeric(database2$No_incu)) 
ggqqplot(as.numeric(database2$No_incu), ylab = "No_incubation")
shapiro.test(as.numeric(database2$F_OSW)) 
ggqqplot(as.numeric(database2$F_OSW), ylab = "Feed_OSW")
shapiro.test(as.numeric(database2$F_SS)) 
ggqqplot(as.numeric(database2$F_SS), ylab = "Feed_SS")
shapiro.test(as.numeric(database2$F_M)) 
ggqqplot(as.numeric(database2$F_M), ylab = "Feed_manure")
shapiro.test(as.numeric(database2$T_constant)) 
ggqqplot(as.numeric(database2$T_constant), ylab = "T_constant")
shapiro.test(as.numeric(database2$T_one_step)) 
ggqqplot(as.numeric(database2$T_one_step), ylab = "T_one_step")
shapiro.test(as.numeric(database2$T_step_wise)) 
ggqqplot(as.numeric(database2$T_step_wise), ylab = "T_stepwise")
shapiro.test(as.numeric(database2$T_i)) 
ggqqplot(as.numeric(database2$T_i), ylab = "T_initial")
shapiro.test(as.numeric(database2$T_f)) 
ggqqplot(as.numeric(database2$T_f), ylab = "T_final")
shapiro.test(as.numeric(database2$OLR_Cons)) 
ggqqplot(as.numeric(database2$OLR_Cons), ylab = "OLR_constant")
shapiro.test(as.numeric(database2$OLR_Stepw)) 
ggqqplot(as.numeric(database2$OLR_Stepw), ylab = "OLR_stepwise")
shapiro.test(as.numeric(database2$OLR_Dyna)) 
ggqqplot(as.numeric(database2$OLR_Dyna), ylab = "OLR_dynamic")
shapiro.test(as.numeric(database2$Lab)) 
ggqqplot(as.numeric(database2$Lab), ylab = "Lab_reactor")
shapiro.test(as.numeric(database2$Pilo_Full)) 
ggqqplot(as.numeric(database2$Pilo_Full), ylab = "Pilot/full_reactor")
shapiro.test(as.numeric(database2$Working.volume)) 
ggqqplot(as.numeric(database2$Working.volume), ylab = "Working.volume")
```

## 5.2. correlation matrix

```{r}
corr <- cor(x = as.matrix(database2[,-c(1,11,12,21:25)]), method = "spearman", use = "pairwise.complete.obs")
corrplot(corr, type="upper", order="hclust", tl.col="black", tl.srt=45)
```

## 5.3. p-value of the important correlations

```{r}

cor.test(database2$F_SS, database2$T_constant, method= "spearman")
cor.test(database2$No_incu, database2$T_one_step, method= "spearman")
cor.test(database2$F_M, database2$T_f, method= "spearman")
cor.test(database2$F_SS, database2$T_f, method= "spearman")
cor.test(database2$F_SS, database2$T_one_step, method= "spearman")

```

# 6. Model-based ranking
## 6.1. random forest

```{r}
set.seed(7789)
better <- tuneRF(database3[, -15], database3[, 15], 
                 stepFactor = 0.5, 
                 plot = T,
                 ntreeTry = 300, 
                 trace = T, 
                 improve = 0.05,
                 ylab = "Number of variables sampled")
randomf_model <- randomForest (Period_t~., 
                                data = database3[,-1], 
                                mtry = 10,
                                ntree = 300,
                                localImp = TRUE,
                                importance = T, proximity = T)
h1 <- hist(treesize(randomf_model),
     xlim= c(17, 30),
     breaks = 9, 
     Main = "Frequency of tree size",
     xlab = "Number of nodes",
     col = "#CCCCCC")
```

## 6.2. Variable importance
### 6.2.1.Minimal depth distribution

```{r}
min_depth <- min_depth_distribution(randomf_model)
save(min_depth, file = "min_depth_frame.rda")
load("min_depth_frame.rda")
head(min_depth, n = 10)
min_deth <-plot_min_depth_distribution(min_depth, mean_sample = "relevant_trees", k = 16, mean_scale = T, mean_round = 3)
```

### 6.2.2. MSE increase afrer permutation and Node purity decrease

```{r}

importance_frame <- measure_importance(randomf_model)
mse_increase <- arrange(importance_frame, desc(mse_increase))
mse_increase$variable <- as.factor(mse_increase$variable)
mse_increase$mse_increase <- as.factor(mse_increase$mse_increase)

mse <- mse_increase %>%
  ggplot(aes(x=mse_increase, y = reorder(variable, mse_increase)))+
  geom_point(shape=23, fill="blue", size=3)+
  ylab("Variable") + xlab("MSE increase")+
  geom_segment(x=0, y= mse_increase$variable, xend = mse_increase$mse_increase, yend =mse_increase$variable,
               color="black", linetype="dashed")


node_purity <- as.data.frame (arrange(importance_frame, desc(node_purity_increase)))
node_purity$variable <- as.factor(node_purity$variable)
node_purity$node_purity_increase <- as.factor (node_purity$node_purity_increase)

node <- node_purity %>%
  ggplot(aes(x=node_purity_increase, y = reorder(variable, node_purity_increase)))+
  geom_point(shape=23, fill="blue", size=3)+
  ylab("Variable") + xlab("Node purity increase")+
  geom_segment(x=0, y= node_purity$variable, xend = node_purity$node_purity_increase, yend =node_purity$variable,
               color="black", linetype="dashed")

ggarrange(min_deth,                                                 
          ggarrange(mse, node, ncol = 2, labels = c("B", "C")), 
          nrow = 2, 
          labels = "A")

```

## 6.3. Partial dependance

```{r}

pred <- predict(randomf_model, database3, probability = TRUE)
table(database3$Period_t, pred)

par.HRT_f <- partial(randomf_model, pred.var = c("HRT_f"), chull = TRUE)
plot.HRT_f <- autoplot(par.HRT_f, contour = TRUE)

par.VS_feed <- partial(randomf_model, pred.var = c("VS_feed"), chull = TRUE)
plot.VS_feed <- autoplot(par.VS_feed, contour = TRUE)

par.TS_feed <- partial(randomf_model, pred.var = c("TS_feed"), chull = TRUE)
plot.TS_feed <- autoplot(par.TS_feed, contour = TRUE)

par.HRT_i <- partial(randomf_model, pred.var = c("HRT_i"), chull = TRUE)
plot.HRT_i <- autoplot(par.HRT_i, contour = TRUE)

par.T_f <- partial(randomf_model, pred.var = c("T_f"), chull = TRUE)
plot.T_f <- autoplot(par.T_f, contour = TRUE)

par.work <- partial(randomf_model, pred.var = c("Working.volume"), chull = TRUE)
plot.work <- autoplot(par.work, contour = TRUE)

par.feed <- partial(randomf_model, pred.var = c("Feed"), chull = TRUE)
plot.feed <- autoplot(par.feed, contour = TRUE)

par.Ti <- partial(randomf_model, pred.var = c("T_i"), chull = TRUE)
plot.Ti <- autoplot(par.Ti, contour = TRUE)

par.Mean_OLR <- partial(randomf_model, pred.var = c("Mean_OLR"), chull = TRUE)
plot.Mean_OLR <- autoplot(par.Mean_OLR, contour = TRUE)

par.Design_OLR <- partial(randomf_model, pred.var = c("Design_OLR"), chull = TRUE)
plot.Design_OLR <- autoplot(par.Design_OLR, contour = TRUE)

par.Incubation <- partial(randomf_model, pred.var = c("Incuba"), chull = TRUE)
plot.Incubation <- autoplot(par.Incubation, contour = TRUE)

par.T_strategy <- partial(randomf_model, pred.var = c("T_strategy"), chull = TRUE)
plot.T_strategy <- autoplot(par.T_strategy, contour = TRUE)

par.Inocu_type <- partial(randomf_model, pred.var = c("Inoc_type"), chull = TRUE)
plot.Inocu_type <- autoplot(par.Inocu_type, contour = TRUE)

par.OLR_strategy <- partial(randomf_model, pred.var = c("OLR_strategy"), chull = TRUE)
plot.OLR_strategy <- autoplot(par.OLR_strategy, contour = TRUE)

par.R_scale <- partial(randomf_model, pred.var = c("R_scale"), chull = TRUE)
plot.R_scale <- autoplot(par.R_scale, contour = TRUE)

ggarrange(plot.HRT_f, plot.VS_feed, plot.TS_feed, plot.HRT_i,plot.T_f, plot.work, plot.Ti,plot.Mean_OLR,plot.Design_OLR,plot.Incubation,plot.T_strategy,
          plot.Inocu_type,plot.OLR_strategy, plot.R_scale, 
          ncol = 3, nrow = 5)

```